Predicting I/O-performance in HPC using Artificial Neural Networks

### 2016-06-25 Paper/template.tex
anfang eval

\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{subcaption}
% ------------

@@

\subsubsection{I/O-path for access time prediction}
Due to the exponential decay of processing speed in the hierarchy, file accesses with similar access parameters, but varying I/O-path, are differentiated with a step in the magnitude of access time.
As the access time is dominated by the slowest component along the I/O-path, a step in the measured access time occurs between to I/O-paths of diverging length.
In figure \ref{src/plot_SizeSorted_log_read_seq.png} measurements of reading file accesses can seen, this figure will be discussed in further detail in the following chapter.
\fig{width=.47\textwidth}{src/plot_SizeSorted_log_read_seq.png}{Measurements of reading file accesses. Access sizes increase from left to right and all measurements with equal access parameter values have the same color.}\\
The different groups of measurements with equal access parameter values (points with the same color) are clearly visible.
Like described previously the different groups are differentiated with a step in the magnitude of access time and can be explained with varying I/O-paths.

@@

The error-class-model is used as described in chapter \ref{sec:error_classes} to examine the potential improvements for access time prediction due to knowledge about I/O-paths.
\section{Evaluation}

\subsection{Test system}

\subsection{Benchmark-tests}

\subsection{Analysis of measurements}
\label{sec:measurements}

>>>>

\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{csvsimple}
% ------------

@@

\subsubsection{I/O-path for access time prediction}
\label{sec:path_for_pred}
Due to the exponential decay of processing speed in the hierarchy, file accesses with similar access parameters, but varying I/O-path, are differentiated with a step in the magnitude of access time.
As the access time is dominated by the slowest component along the I/O-path, a step in the measured access time occurs between to I/O-paths of diverging length.
In figure \ref{src/plot_SizeSorted_log_read_seq.png} measurements of reading file accesses can seen, this figure will be discussed in further detail in the following chapter.
\fig{width=.47\textwidth}{src/plot_SizeSorted_log_read_seq.png}{Measurements of reading file accesses. Access sizes increase from left to right and all measurements with equal access parameter values have the same color.}\\
The different groups of measurements with equal access parameter values (points with the same color) are clearly visible.
Like described previously the different groups are differentiated with a step in the magnitude of access time and can be explained with varying I/O-paths.

@@

The error-class-model is used as described in chapter \ref{sec:error_classes} to examine the potential improvements for access time prediction due to knowledge about I/O-paths.

\section{Evaluation}
The execution and results of analysis are summarized in the following.
First the test system used for measurements and the strategy for systematic measurement are presented.
Then follows the study of received data and computed error classes.
And in the end the quality of predictions of the different models are examined.

\subsection{Test system}
The measurements were done on the super computer Mistral of the DKRZ (Deutsches Klimarechenzentrum).
Mistral operates with the parallel distributed file system Lustre and consists of over 1500 computing nodes and 30 petabyte storage with a memory bandwidth of 300\,GiB/s.
Each computing node consists of two E5-2680v3 with a clock rate of 2.5\,GHz and 30\,MiB L3 Cache.
Measurements were done during daily operation, typical fluctuations in workload may have influenced the execution of benchmarks.

\subsection{Benchmarking}
We are examining the test system with sequences of measurements.
Each series of measurements follows a certain pattern of file access to study system behavior systematically.\\
Two different use cases were tested, each with reading and writing file accesses.
On the one hand sequential file access and on the other random file access.
In the sequential case data directly following the previous access is retrieved.
The random pattern is accessing the file at arbitrary positions.\\
Our test file has a size of 10\,GiB which doesn't fit in the main memory of computer nodes.
Occasional access of hard drives in the parallel storage system is likely.
Most series of measurements contain 10\,000 file accesses with a certain access size, only series with sequential access and an access size of greater then 2\,MiB have fewer measurements as the end of our 10\,GiB file is reached.
Every series with specific access pattern, access size and access type is repeated three times.
Access sizes are varying from 1\,B to 16\,MiB (in detail:  1\,B, 4\,B, 16\,B, 64\,B, 256\,B, 1\,KiB, 4\,KiB, 8\,KiB, 16\,KiB, 64\,KiB, 256\,KiB, 512\,KiB, 1\,MiB, 2\,MiB, 4\,MiB, 8\,MiB and 16\,MiB).

\subsection{Analysis of measurements}
\label{sec:measurements}
At first the measured access times for the four different cases are summarized in table \ref{overview}.
\tab{overview}{Overview on file access times for the different cases.}{
	\centering
	\scriptsize
	\begin{tabular}{|r|r|r|r|r|r|r|}\hline%
		Case & Min. value  & 1. Quartile & Median & Arith. mean & 3. Quartile & Max. value \\\hline\hline
		\csvreader[late after line=\\\hline]%
		{src/data_summary.csv}{Attribut=\Attribut,Min=\Min,Quartil1=\L, Median = \Median, Mittel = \Mittel,Quartil3 = \Q, Max = \Max, Korrelation = \Korrelation}%
		{\Attribut & \Min & \L & \Median & \Mittel & \Q & \Max}%
	\end{tabular}
}\\
One can find that sequential file accesses are on average much faster than random accesses especially for the reading case.
Interestingly our measured random reading file accesses were slower than writing reading accesses.\\
In figure \ref{correlations} we consider the correlations between file access parameters and the resulting access time, a strong correlation can be exploited for access time prediction.
\tab{correlations}{Correlations between file access parameters and access time.}{
	\centering
	\scriptsize
	\begin{tabular}{|r|r|r|}\hline%
		Attribute & Sequential access pattern & Random access pattern\\\hline\hline
		\csvreader[late after line=\\\hline]%
		{src/data_summary_korrelationen.csv}{Attribut=\Attribut,seq = \seq, rnd = \rnd}%
		{\Attribut & \seq &\rnd}%
	\end{tabular}
}\\
It is to note that sequential access leads to a constant delta-offset of $0$.
Sequential file accesses have with 97.3\% a clear correlation to access time, this is with only 32.91\% to a lesser degree still true for the random accesses.
This is why linear regression can be expected to have acceptable results for access time prediction, especially for the sequential access case.\medskip

To study the distribution of measured file access times we display them as follows:
All measurements with equal access size have the same color, access sizes of file accesses are increasing in the graphs from left to right.
The resulting graphs can be seen in the figures \ref{src/plot_SizeSorted_log_read_seq2.png} to \ref{src/plot_SizeSorted_log_write_rnd.png}.
\fig{width=.47\textwidth}{src/plot_SizeSorted_log_read_seq2.png}{Measurements of sequential reading file accesses.}
\fig{width=.47\textwidth}{src/plot_SizeSorted_log_write_seq.png}{Measurements of sequential writing file accesses.}
\fig{width=.47\textwidth}{src/plot_SizeSorted_log_read_rnd.png}{Measurements of random reading file accesses.}
\fig{width=.47\textwidth}{src/plot_SizeSorted_log_write_rnd.png}{Measurements of random writing file accesses.}
For clarity the slowest and fastest 1\% of measurements are cut and only every 25th measurement is represented.\\
The correlation between access size and access time can be easily identified.
As described in chapter \ref{sec:path_for_pred} the split of measurements with equal access parameter values, that occurs in particular in the cases of sequential file accesses, can be explained with a processing in the storage system along different I/O-paths.
Worth mentioning is that the groups of access time for different access sizes are partially at the same range of magnitude.
This indicates as well that the splits are caused by varying I/O-paths as one I/O-path has a typical access time, which only changes slightly for different access sizes.\\
The phenomenon of these splits clarifies why despite of the strong correlation of access size to access time a linear model might produce sub-optimal predictions. It's difficult to predict which I/O-path will be used for a measurement as they differentiate for equal access parameter values.
In figure \ref{src/plot_SizeSorted_log_read_rnd.png} the importance of knowledge about I/O-paths for the  access time prediction can be seen, for example file accesses of only 1\,B can be processed as slow as accesses with 16\,MiB if the longest I/O-path is used.\\
For writing file accesses the splits only occur occasionally and are less drastic; these measurements might be well represented with a linear model.
